Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2023-12-21 doc/manuscript/translocation_bioarxiv.tex
Minor edits to Intro and Results

\dropcap{H}uman activities are increasingly impacting global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. One consequence of human
alteration of the biosphere is an increase in emerging infectious
diseases \citep{jones2008, fisher2012}. Such diseases pose a severe
threat to wildlife populations \citep{daszak2000}, and have caused
dramatic declines and extinctions in a wide range of taxa, including
echinoderms, mammals, birds, and amphibians
\citep{hewson2014, samuel2015, scheele2019, cunningham2021}. Amphibians
are experiencing particularly devastating impacts of disease due to the
recent emergence and global spread of the highly virulent amphibian
chytrid fungus, \emph{Batrachochytrium dendrobatidis} (Bd)
\citep{luedtke2023, scheele2019}. By one estimate, hundreds of species
have experienced Bd-caused declines, and numerous susceptible taxa are
extinct in the wild \citep{scheele2019}. These impacts to global
biodiversity may be unprecedented, and highlight the importance of
understanding mechanisms of species persistence in the presence of
emerging diseases \citep{russell2020}.

Evidence of natural recovery in the many Bd-impacted amphibian
populations is surprisingly rare \citep[for notable exceptions,
see][]{scheele2017, voyles2018, knapp2016}, suggesting that
disease-caused declines will be difficult to reverse. This apparent low
resilience to disease effects may be due to the limited ability of many
amphibians to develop Bd resistance and/or tolerance, which in turn,
could also lessen the effectiveness of potential Bd mitigation
strategies. Following pathogen arrival in a host population, resistance
(ability to limit pathogen burden) and tolerance (ability to limit the
harm caused by a particular burden) are key mechanisms to reduce disease
impacts \citep{raberg2009} and facilitate population persistence and
recovery \citep{brannelly2021}. Host immunity and evolution both play
important roles in the development of resistance and tolerance, and
utilizing these factors would seem a promising approach to developing
effective strategies to mitigate disease impacts in the wild
\citep{garner2016, woodhams2011}. However, several aspects of the
amphibian-Bd system present difficult obstacles, including (i) the
general inability of amphibians to mount an effective immune response
against Bd infection \citep{rosenblum2012, fites2013, grogan2018a}, and
(ii) the apparent rarity of evolution of more resistant/tolerant
genotypes \citep[but see][]{savage2016, grogan2018b}. These factors
suggest that reintroduction of amphibians into sites to reestablish
populations extirpated by Bd will often result not in population
recovery, but instead in the rapid reinfection and mortality of the
introduced animals and/or their progeny
\citep{hammond2021, knapp2022, stockwell2008, scheele2021}. If true, the
future of many amphibian species threatened by Bd appears bleak.

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the global declines of amphibians caused by Bd
\citep{scheele2019}. Once the most common amphibian in the high
elevation portion of California's Sierra Nevada mountains
\citep[USA,][]{grinnell1924}, during the past century this frog has
disappeared from more than 90\% of its historical range
\citep{vredenburg2007}. Due to the severity of its decline and the
increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000} starting in the late 1800s. The arrival
of Bd in the mid-1900s and its subsequent spread \citep{vredenburg2019}
caused additional large-scale population extirpations
\citep{vredenburg2010, rachowicz2006}. These Bd-caused declines are
fundamentally different from the fish-caused declines because fish
eradication is feasible \citep{knapp1998} and results in the rapid
recovery of frog populations \citep{knapp2007, vredenburg2004}. In
contrast, Bd appears to persist in habitats even in the absence of
amphibian hosts \citep{walker2007}, and therefore represents a long-term
alteration of invaded ecosystems that amphibians will need to overcome
to reestablish populations.

Despite the catastrophic impact of Bd on MYL frogs, wherein most
Bd-naive populations are extirpated following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep[during which Bd infection intensity on frogs is very
high,][]{briggs2010} and are now recovering
(Figure~\ref{fig-recovery-model}) \citep{knapp2016}. Frogs in these
recovering populations show reduced susceptibility to Bd infection
\citep{knapp2016}, with infection intensity (``load'') on adults
consistently in the low-to-moderate range
\citep{briggs2010, knapp2011, joseph2018}. This reduced susceptibility
is evident even under controlled laboratory conditions
\citep{knapp2016}, indicative of host resistance against Bd infection
(and not simply an effect of factors external to individual frogs, e.g.,
environmental conditions). In addition to frogs from recovering
populations having higher resistance to Bd infection than those from
naive populations, they could also have higher tolerance, but no data
are currently available to evaluate this possibility. Therefore, we
focus on resistance throughout this paper. The observed resistance of
MYL frogs could be the result of several non-mutually exclusive
mechanisms, including natural selection for more resistant genotypes
\citep{savage2016, grogan2018b}, acquired immunity \citep{grogan2018a},
and/or inherent between-population differences that pre-date Bd
exposure. The possible evolution of MYL frog resistance and subsequent
population recovery is consistent with that expected under
``evolutionary rescue'', whereby rapid evolutionary change increases the
frequency of adaptive alleles and restores positive population growth
\citep{carlson2014, searle2020}. This intriguing possibility also
suggests an opportunity to expand recovery beyond the spatial scale
possible under natural recovery by utilizing resistant frogs from
recovering populations in reintroductions to vacant habitats
(Figure~\ref{fig-recovery-model}) \citep{joseph2018, mendelson2019}.

In the current study, we had three primary objectives. First, to
determine whether the reintroduction of resistant MYL frogs obtained
from populations recovering from Bd-caused declines allows the
successful reestablishment of extirpated populations despite ongoing
disease, we conducted a 15-year landscape-scale frog reintroduction
effort (Figure~\ref{fig-recovery-model}). Second, to extend our
inferences of population recovery well beyond the temporal extent of our
reintroduction study, we developed a model to estimate the probability
of persistence for the reintroduced populations over a multi-decadal
period (Figure~\ref{fig-recovery-model}). Third, we investigated whether
MYL frogs show evidence of an evolutionary response to Bd exposure, and
whether these genomic changes are associated with resistance. To
accomplish this objective, we used exome capture methods to identify
genomic regions with consistent differences between frogs from naive
versus recovering populations (Figure~\ref{fig-recovery-model}) and
determined whether any of these regions contain genes associated with
resistance.

@@

Bd loads were fairly consistent before versus after translocation, and
loads were nearly always well below the level indicative of severe
chytridiomycosis (i.e., the disease caused by Bd) and associated frog
mortality (Figure S2)
\citep{joseph2018, vredenburg2010}. Although it is possible that the
observed relatively small changes in load are a consequence of
individuals with high Bd loads dying and therefore being unavailable for
sampling during the post-translocation period, the fact that there was
little difference in pre- versus post-translocation Bd loads even in
those populations that had very high frog survival (70556, 74976;
Figure~\ref{fig-translocation-survival},
Figure S2) suggests a true lack of
substantial change in Bd load.

The ultimate measure of reintroduction success is the establishment of a
self-sustaining population. Given that it can take years or even decades
to determine the self-sustainability of a reintroduced population
\citep[for an example in MYL frogs, see][]{joseph2018}, the use of
proxies is essential for providing shorter-term insights into
reintroduction success and the factors driving it. Results from our CMR
surveys allowed us to accurately estimate frog survival, including over
the entire CMR time series for each site and during only the 1-year
period immediately following translocation. These estimates were made
using site-specific models analyzed using the mrmr package. We use these
estimates to describe general patterns of frog survival in all
translocated cohorts, and in an among-site meta-analysis of frog
survival to identify important predictors of 1-year frog survival (e.g.,
Bd load).

@@

The goal of our meta-analysis was to identify important predictors of
1-year frog survival. We were particularly interested in whether Bd load
had a negative effect on adult survival, as would be expected if frogs
were highly susceptible to Bd infection. This analysis was conducted in
a Bayesian framework and included a diversity of site, cohort, and
individual level characteristics as predictors and 1-year frog survival
(Figure~\ref{fig-translocation-survival}) as the response variable. The
best model of 1-year frog survival identified several important
predictors, but Bd load at the time of translocation was not among them
(Figure S3). Instead, important predictors
included winter severity in the year following translocation (snow\_t1),
site elevation, and donor population (Figure~\ref{fig-cond-effects},
Figure S3). Males had somewhat higher
survival than females, but this effect was small
(Figure~\ref{fig-cond-effects}, Figure S3).
The absence of Bd load as an important predictor of frog survival is
consistent with frogs in recovering populations having sufficient
resistance to suppress Bd loads below harmful levels.

@@

Results from the frog translocation study indicated that most
populations showed evidence of successful reproduction and recruitment,
and that adult survival was often relatively high (described above).
Although suggestive of population establishment, a decade or more of
surveys may be necessary to confirm that populations are in fact
self-sustaining \citep{joseph2018}. To extend our inferences of
population establishment beyond those possible from the site-specific
CMR data, we developed a population viability model. Specifically, to
test whether the observed yearly adult survival probabilities in
translocated populations (provided in legend of
Figure~\ref{fig-viability}B) were sufficient for long-term viability,
we built a stage-structured matrix model that captured known frog
demography and included demographic and environmental stochasticity. We
parameterized the model using CMR data from translocated populations and
known life history values in this system (Table S2
SI).

@@

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in year-1 juvenile survival and recruitment
(the transition that we expect to be the most subject to environmental
variability in the presence of Bd), populations with high adult survival
are likely to persist over a 50 year time horizon. Our model predicted
that, following a single introduction of 40 adult individuals into a
population, the six populations with the highest adult survival
probabilities (\(\sigma_{A_R} > 0.5\)) had 50-year extinction
probabilities of less than 0.5 when the average year 1 juvenile survival
was greater than 0.10 (Figure~\ref{fig-viability}B). This indicates
strong potential for long-term persistence in the presence of Bd and
environmental variability in survival and recruitment. In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be \textgreater{} 50\% regardless of the value of mean year
1 juvenile survival between 0 and 0.25. To test the validity of our
model predictions, we demonstrated that our stochastic model could
describe the general recovery trajectory of our translocated population
with the longest survey history (Figure~\ref{fig-viability}C;
population 70550, surveyed for 16 years).

@@

To determine whether MYL frog populations show an evolutionary response
to Bd, we compared frog exomes (i.e., coding region of genome) between
populations with contrasting histories of Bd exposure. Specifically, we
compared frog genomes sampled in 4 populations that have not yet
experienced a Bd-caused epizootic (``naive'') \citep{zhou2015} versus in
5 populations that experienced a Bd epizootic during the past several
decades and have since recovered to varying degrees (``recovering'';
Figure~\ref{fig-allelefrequencies}) \citep{knapp2016, vredenburg2010}.
Bd-exposure histories of the 9 study populations are based on 10-20
years of VES and Bd surveillance using skin swabbing
\citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive populations are
characterized by large numbers of adults (i.e, typically 1000s), Bd
prevalence that is generally 0\% except during occasional Bd failed
invasions \citep[during which Bd loads remain very low,][]{wilber2022},
and no history of Bd epizootics since we first surveyed these
populations in the late 1990s and early 2000s \citep{zhou2015}. In
contrast, recovering populations exist in an enzootic state
\citep{briggs2010}, characterized by smaller numbers of adults
(generally \textless{} 100), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation. (See \textbf{Supporting Information - Frog evolution in
response to Bd - Methods} for additional details regarding the study
design.)

We conducted a principal component analysis (PCA) of the genomic data to
describe the relationships between sampled populations, and then used
two complementary approaches to identify regions of the genome that
differed between naive and recovering populations (i.e., regions
putatively under selection). First, we used a multivariate linear mixed
model to evaluate associations between population type (i.e., naive
versus recovering) and individual variants, including single nucleotide
polymorphisms (SNPs) and insertions/deletions (INDELS), while accounting
for population structure. Second, we used a splined window analysis to
identify larger genomic regions showing differences between population
types in \emph{F\textsubscript{ST}} and nucleotide diversity
(\(\pi_{diff}\) = \(\pi_{naive}\) - \(\pi_{recovering}\)).

Individual frogs clustered into 3 separate groups in principal component
space (Figure S6A SI), and clusters reflected
the species split (i.e., \emph{R. muscosa} versus \emph{R. sierrae}) and
the strong signature of isolation by distance that is characteristic of
MYL frogs \citep{rothstein2020, poorten2017, byrne2023}. Importantly,
each cluster contained at least one population from both the naive and
recovering groups, allowing us to distinguish allelic associations of
individuals sampled in the 2 population types versus allelic
associations resulting from population structure and genetic drift.

@@

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 33 annotated genes were found in the 9 shared regions
(Datasets S3, S4). Given this large number of genes, here we focus on
those with the strongest signal of selection and/or immune-related
functions. The largest \(\pi_{diff}\), indicative of directional
selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of
the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan}C). This region
contains approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

No gene ontology (GO) biological functions, molecular functions, or
cellular processes were over-represented in either the outlier variants
or the 33 genes located in the overlapping \emph{F\textsubscript{ST}}
and \(\pi_{diff}\) splined windows. In addition, although the joint
processes of Bd-caused population declines and selection in response to
Bd exposure could affect genetic diversity of recovering populations, we
found no consistent differences in individual-level heterozygosity or
population-level \(\pi\) between naive and recovering populations (see
\textbf{Supporting Information-Frog evolution in response to Bd-Methods}
for details).

In summary, these results indicate that the exomes of frogs from our
naive and recovering study populations show substantial differences,
consistent with parallel evolutionary changes following Bd exposure. The
regions putatively under selection contain several
immunologically-relevant genes and gene families that are directly
linked to disease resistance in other taxa.

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but broadly-applicable strategies to recover
affected species are generally lacking \citep[e.g.,][]{garner2016}. Our
results indicate (i) the capacity of reintroduced populations to become
established and eventually recover despite ongoing disease, (ii) that
the recovering populations are likely to persist over a 50-year period,
(iii) that there are substantial genomic differences between naive and
recovering MYL frog populations, consistent with evolutionary change in
frogs following Bd exposure, and (iv) that some of the genomic regions
putatively under selection contain genes related to disease resistance.
Collectively, these results (Figure~\ref{fig-recovery-model}) provide a
rare example of amphibian recovery in the presence of a highly virulent
pathogen, and have important implications for the conservation and
recovery of amphibians and other taxa worldwide that are endangered by
escalating impacts from emerging infectious diseases. In light of the
generally low success rate of amphibian reintroduction efforts
\citep{dodd2005}, our success in reestablishing MYL frog populations via
translocation of resistant individuals is striking, and even more so
given that this success occurred despite ongoing infection of frogs with
Bd.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also Figure S2]{briggs2010, knapp2011}. The differences in Bd load of frogs from
naive and recovering populations are also observed in controlled
laboratory studies \citep[see Figure 4 in][]{knapp2016}, and clearly
indicate that frogs from recovering populations exhibit resistance
against Bd infection. This resistance could in theory be due to several
factors, including natural selection for more resistant genotypes,
acquired immunity, and/or inherent between-population differences that
pre-date Bd exposure, but until now evidence to evaluate the role of
evolution was lacking.

>>>>

\dropcap{H}uman activities are increasingly impacting global biodiversity (1),
with important implications for ecosystem resilience and human welfare
(2). One consequence of human alteration of the biosphere is an increase
in emerging infectious diseases (3, 4). Such diseases pose a severe
threat to wildlife populations (5), and have caused dramatic declines
and extinctions in a wide range of taxa, including echinoderms, mammals,
birds, and amphibians (6--9). Amphibians are experiencing particularly
devastating impacts of disease due to the recent emergence and global
spread of the highly virulent amphibian chytrid fungus,
\emph{Batrachochytrium dendrobatidis} (Bd) (8, 10). By one estimate,
hundreds of species have experienced Bd-caused declines, and numerous
susceptible taxa are extinct in the wild (8). These impacts to global
biodiversity may be unprecedented, and highlight the importance of
understanding mechanisms of species persistence in the presence of
emerging diseases (11).

Evidence of natural recovery in the many Bd-impacted amphibian
populations is surprisingly rare (for notable exceptions, see 12, 13,
14), suggesting that disease-caused declines will be difficult to
reverse. This apparent low resilience to disease effects may be due to
the limited ability of many amphibians to develop Bd resistance and/or
tolerance, which in turn, could also lessen the effectiveness of
potential Bd mitigation strategies. Following pathogen arrival in a host
population, resistance (ability to limit pathogen burden) and tolerance
(ability to limit the harm caused by a particular burden) are key
mechanisms to reduce disease impacts (15) and facilitate population
persistence and recovery (16). Host immunity and evolution both play
important roles in the development of resistance and tolerance, and
utilizing these factors would seem a promising approach to developing
effective strategies to mitigate disease impacts in the wild (17, 18).
However, several aspects of the amphibian-Bd system present difficult
obstacles, including (i) the general inability of amphibians to mount an
effective immune response against Bd infection (19--21), and (ii) the
apparent rarity of evolution of more resistant/tolerant genotypes (but
see 22, 23). These factors suggest that reintroduction of amphibians
into sites to reestablish populations extirpated by Bd will often result
not in population recovery, but instead in the rapid reinfection and
mortality of the introduced animals and/or their progeny (24--27). If
true, the future of many amphibian species threatened by Bd appears
bleak.

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} (28), is emblematic of the
global declines of amphibians caused by Bd (8). Once the most common
amphibian in the high elevation portion of California's Sierra Nevada
mountains (USA, 29), during the past century this frog has disappeared
from more than 90\% of its historical range (28). Due to the severity of
its decline and the increasing probability of extinction, both species
are now listed as ``endangered'' under the U.S. Endangered Species Act.
In the Sierra Nevada, this decline was initiated by the introduction of
non-native trout into the extensive historically-fishless region (30,
31) starting in the late 1800s. The arrival of Bd in the mid-1900s and
its subsequent spread (32) caused additional large-scale population
extirpations (33, 34). These Bd-caused declines are fundamentally
different from the fish-caused declines because fish eradication is
feasible (35) and results in the rapid recovery of frog populations (36,
37). In contrast, Bd appears to persist in habitats even in the absence
of amphibian hosts (38), and therefore represents a long-term alteration
of invaded ecosystems that amphibians will need to overcome to
reestablish populations.

Despite the catastrophic impact of Bd on MYL frogs, wherein most
Bd-naive populations are extirpated following Bd arrival (33), some
populations have persisted after epizootics (during which Bd infection
intensity on frogs is very high, 39) and are now recovering
(Figure~\ref{fig-recovery-model}) (14). Frogs in these recovering
populations show reduced susceptibility to Bd infection (14), with
infection intensity (``load'') on adults consistently in the
low-to-moderate range (39--41). This reduced susceptibility is evident
even under controlled laboratory conditions (14), indicative of host
resistance against Bd infection (and not simply an effect of factors
external to individual frogs, e.g., environmental conditions). In
addition to frogs from recovering populations having higher resistance
to Bd infection than those from naive populations, they could also have
higher tolerance, but no data are currently available to evaluate this
possibility. Therefore, we focus on resistance throughout this paper.
The observed resistance of MYL frogs could be the result of several
non-mutually exclusive mechanisms, including natural selection for more
resistant genotypes (22, 23), acquired immunity (21), and/or inherent
between-population differences that pre-date Bd exposure. The possible
evolution of MYL frog resistance and subsequent population recovery is
consistent with that expected under ``evolutionary rescue'', whereby
rapid evolutionary change increases the frequency of adaptive alleles
and restores positive population growth (42, 43). This intriguing
possibility also suggests an opportunity to expand recovery beyond the
spatial scale possible under natural recovery by utilizing resistant
frogs from recovering populations in reintroductions to vacant habitats
(Figure~\ref{fig-recovery-model}) (41, 44).

In the current study, we had three primary objectives. First, to
determine whether the reintroduction of resistant MYL frogs obtained
from populations recovering from Bd-caused declines allows the
successful reestablishment of extirpated populations despite ongoing
disease, we conducted a 15-year landscape-scale frog reintroduction
effort (Figure~\ref{fig-recovery-model}). Second, to extend our
inferences of population recovery well beyond the temporal extent of our
reintroduction study, we developed a model to estimate the probability
of persistence for the reintroduced populations over a multi-decadal
period (Figure~\ref{fig-recovery-model}). Third, given the importance of
resistance for frog survival, population establishment, and long-term
viability (this study), we also conducted a genomic study to determine
whether MYL frogs show an evolutionary response to Bd exposure, and
whether these genomic changes are associated with resistance. To
accomplish this objective, we used exome capture methods to identify
genomic regions with consistent differences between frogs from naive
versus recovering populations (Figure~\ref{fig-recovery-model}).

@@

Bd loads were fairly consistent before versus after translocation, and
loads were nearly always well below the level indicative of severe
chytridiomycosis (i.e., the disease caused by Bd) and associated frog
mortality (Figure S2) (33, 41). Although it
is possible that the observed relatively small changes in load are a
consequence of individuals with high Bd loads dying and therefore being
unavailable for sampling during the post-translocation period, the fact
that there was little difference in pre- versus post-translocation Bd
loads even in those populations that had very high frog survival (70556,
74976; Figure~\ref{fig-translocation-survival},
Figure S2) suggests a true lack of
substantial change in Bd load.

The ultimate measure of reintroduction success is the establishment of a
self-sustaining population. Given that it can take years or even decades
to determine the self-sustainability of a reintroduced population (for
an example in MYL frogs, see 41), the use of proxies is essential for
providing shorter-term insights into reintroduction success and the
factors driving it. Results from our CMR surveys allowed us to
accurately estimate frog survival, including over the entire CMR time
series for each site and during only the 1-year period immediately
following translocation. These estimates were made using site-specific
models analyzed using the mrmr package. We use these estimates to
describe general patterns of frog survival in all translocated cohorts,
and in an among-site meta-analysis of frog survival to identify
important predictors of 1-year frog survival (e.g., Bd load).

@@

The goal of our meta-analysis was to identify important predictors of
1-year frog survival. We were particularly interested in whether Bd load
had a negative effect on adult survival, as would be expected if frogs
were highly susceptible to Bd infection. This analysis was conducted in
a Bayesian framework and included a diversity of site, cohort, and
individual- level characteristics as predictors and 1-year frog survival
(Figure~\ref{fig-translocation-survival}) as the response variable. The
best model of 1-year frog survival identified several important
predictors, but Bd load at the time of translocation was not among them
(Figure S3). Instead, important predictors
included winter severity in the year following translocation (snow\_t1),
site elevation, and donor population (Figure~\ref{fig-cond-effects},
Figure S3). Males had somewhat higher
survival than females, but this effect was small
(Figure~\ref{fig-cond-effects}, Figure S3).
The absence of Bd load as an important predictor of frog survival is
consistent with frogs in recovering populations having sufficient
resistance to suppress Bd loads below harmful levels.

@@

Results from the frog translocation study indicated that most
populations showed evidence of successful reproduction and recruitment,
and that adult survival was often relatively high (described above).
Although suggestive of population establishment, a decade or more of
surveys may be necessary to confirm that populations are in fact
self-sustaining (41). To extend our inferences of population
establishment beyond those possible from the site-specific CMR data, we
developed a population viability model. Specifically, to test whether
the observed yearly adult survival probabilities in translocated
populations (provided in legend of Figure~\ref{fig-viability}B) were
sufficient for long-term viability, we built a stage-structured matrix
model that captured known frog demography and included demographic and
environmental stochasticity. We parameterized the model using CMR data
from translocated populations and known life history values in this
system (Table S2 SI).

@@

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in year-1 juvenile survival and recruitment
(the transition that we expect to be the most subject to environmental
variability in the presence of Bd), populations with high adult survival
are likely to persist over a 50 year time horizon. Our model predicted
that, following a single introduction of 40 adult individuals into a
population, the six populations with the highest adult survival
probabilities (\(\sigma_{A_R} > 0.5\)) had 50-year extinction
probabilities of less than 0.5 when the average year-1 juvenile survival
was greater than 0.10 (Figure~\ref{fig-viability}B). This indicates
strong potential for long-term persistence in the presence of Bd and
environmental variability in survival and recruitment. In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be \textgreater{} 50\% regardless of the value of mean
year-1 juvenile survival between 0 and 0.25. To test the validity of our
model predictions, we demonstrated that our stochastic model could
describe the general recovery trajectory of our translocated population
with the longest survey history (Figure~\ref{fig-viability}C;
population 70550, surveyed for 16 years).

@@

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') (45) versus in 5 populations that experienced a Bd epizootic
during the past several decades and have since recovered to varying
degrees (``recovering''; Figure~\ref{fig-allelefrequencies}) (14, 33).
Bd-exposure histories of the 9 study populations are based on 10-20
years of VES and Bd surveillance using skin swabbing (e.g., 14, 45, 46).
Naive populations are characterized by large numbers of adults (i.e,
typically 1000s), Bd prevalence that is generally 0\% except during
occasional Bd failed invasions (during which Bd loads remain very low,
46), and no history of Bd epizootics since we first surveyed these
populations in the late 1990s and early 2000s (45). In contrast,
recovering populations exist in an enzootic state (39), characterized by
smaller numbers of adults (generally \textless{} 100), high Bd
prevalence (often \textgreater{} 80\%, 40), and, in adults, moderate Bd
loads that are typically well below the level expected to cause
mortality (33). Naive and recovering populations can be identified
unambiguously using these differences in Bd prevalence and load.
Finally, there is no potential for frog dispersal between the 9 study
populations due to intervening distances and topography, as well as the
presence of introduced (predatory) fish and fish-induced habitat
fragmentation. (See \textbf{Supporting Information - Frog evolution in
response to Bd - Methods} for additional details regarding the study
design.)

We conducted a principal component analysis (PCA) of the genomic data to
describe the relationships between sampled populations, and then used
two complementary approaches to identify regions of the genome that
differed between naive and recovering populations (i.e., regions under
selection). First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS), while accounting for population
structure. Second, we used a splined window analysis to identify larger
genomic regions showing differences between population types in
\emph{F\textsubscript{ST}} and nucleotide diversity (\(\pi_{diff}\) =
\(\pi_{naive}\) - \(\pi_{recovering}\)).

Individual frogs clustered into 3 separate groups in principal component
space (Figure S6A SI), and clusters reflected
the species split (i.e., \emph{R. muscosa} versus \emph{R. sierrae}) and
the strong signature of isolation-by-distance that is characteristic of
MYL frogs (47--49). Importantly, each cluster contained at least one
population from both the naive and recovering groups, allowing us to
distinguish allelic associations of individuals sampled in the 2
population types versus allelic associations resulting from population
structure and genetic drift.

@@

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 33 annotated genes were found in the 9 shared regions
(Datasets S3, S4). Given this large number of genes, here we focus on
those with the strongest signal of selection and/or immune-related
functions. The largest \(\pi_{diff}\), indicative of directional
selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of
the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan}C). This region
contains approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a large negative
\(\pi_{diff}\) value, consistent with balancing selection. Finally, one
shared outlier region on Contig8 contained one outlier SNP (TCF19) and
was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S7).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) (50). Therefore, this
region under selection is part of the extended MHC Class I and III
complex and shows synteny with other amphibian genomes.

Although the joint processes of Bd-caused population declines and
selection in response to Bd exposure could affect genetic diversity of
recovering populations, we found no consistent differences in
individual-level heterozygosity or population-level \(\pi\) between
naive and recovering populations (see \textbf{Supporting
Information-Frog evolution in response to Bd} for details). Thus,
despite localized selection in particular regions of the genome, we did
not find evidence for reduced genetic diversity across the genome in
recovering populations. In addition, no gene ontology (GO) biological
functions, molecular functions, or cellular processes were
over-represented in either the outlier variants or the 33 genes located
in the overlapping \emph{F\textsubscript{ST}} and \(\pi_{diff}\) splined
windows (see \textbf{Supporting Information-Frog evolution in response
to Bd} for details).

In summary, our genomic results indicate that the exomes of frogs from
naive and recovering study populations show substantial differences,
consistent with parallel evolutionary changes following Bd exposure. The
regions under selection contain several immunologically-relevant genes
and gene families that are directly linked to disease resistance in
other taxa.

@@

Disease-induced population declines are decimating global biodiversity
(5), but broadly-applicable strategies to recover affected species are
generally lacking (e.g., 17). Here, we tested the possibility that
populations of resistant individuals from naturally recovering
populations can be used to reestablish extirpated populations of the
endangered MYL frog in the presence of a highly virulent fungal pathogen
(Bd). Our results indicate (i) the capacity of reintroduced populations
to become established and eventually recover despite ongoing disease,
(ii) that the recovering populations are likely to persist over a
50-year period, (iii) that there are substantial genomic differences
between naive and recovering MYL frog populations, consistent with
evolutionary change in frogs following Bd exposure, and (iv) that some
of the genomic regions under selection contain genes related to disease
resistance. Collectively, these results
(Figure~\ref{fig-recovery-model}) provide a rare example of amphibian
recovery in the presence of Bd, and have important implications for the
conservation and recovery of amphibians and other taxa worldwide that
are endangered by escalating impacts from emerging infectious diseases.
In light of the generally low success rate of amphibian reintroduction
efforts (51), our success in reestablishing MYL frog populations via
translocation of resistant individuals is striking, and even more so
given that this success occurred despite ongoing infection of frogs with
Bd.

In the following discussion, we follow the sequence of frog recovery
described in Figure~\ref{fig-recovery-model} to structure our key
points. Previous field studies in MYL frogs show that frog-Bd dynamics
and frog survival in the presence of Bd are fundamentally different
between naive and recovering populations. Following the arrival and
establishment of Bd in previously-naive populations, adult frogs develop
high Bd loads that lead to mass die-offs (33). In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed (39, 40,
see also Figure S2). The differences in Bd
load of frogs from naive and recovering populations are also observed in
controlled laboratory studies (see Figure 4 in 14), and clearly indicate
that frogs from recovering populations exhibit resistance against Bd
infection. This resistance could in theory be due to several factors,
including natural selection for more resistant genotypes, acquired
immunity, and/or inherent between-population differences that pre-date
Bd exposure, but until now evidence to evaluate the role of evolution
was lacking.

Results from our genomic analyses suggest that natural selection for
adaptive alleles is at least partially responsible for the increased
resistance of frogs in recovering populations. We identified multiple
specific alleles and genomic regions showing signatures of selection
between adjacent naive and recovering MYL frog populations, consistent
with selection following Bd exposure. These analysis are based on
samples collected from virtually all of the MYL frog populations
remaining in a naive state, as well as adjacent recovering populations.
This produced genetic clusters that contained at least one naive and one
recovering population, allowing us to detect selection without the
confounding effects of population structure. Finally, we did not find a
reduction in overall genetic variation in the recovering populations,
suggesting that despite localized selection in the genome, these
populations retain adequate genetic diversity for long-term persistence.
